---
output:
  pdf_document: default
  html_document: default
---
# Bellabeat Case Study

## 1. Scenario
Bellabeat is a high-tech manufacturer of health-focused products for women. Bellabeat is a successful small company, but they have the potential to become a larger player in the global smart device market. Urška Sršen, cofounder and Chief Creative Officer of Bellabeat, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company.

## 2. Ask 
1. What are some trends in smart device usage?
2. How could these trends apply to Bellabeat customers?
3. How could these trends help influence Bellabeat marketing strategy?R

## 3. Prepare: 

#### 3.1 Loading packages
```{r}
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(skimr)
library(janitor)
```

#### 3.2 import data
```{r }
daily_activity <- read_csv("dailyActivity_merged.csv")
daily_calories <- read_csv("dailyCalories_merged.csv")
daily_intensities <- read_csv("dailyIntensities_merged.csv")
daily_steps <- read_csv("dailySteps_merged.csv")
heartrate <- read_csv("heartrate_seconds_merged.csv")
hourly_calories <- read.csv("hourlyCalories_merged.csv")
hourly_intensities <- read_csv("hourlyIntensities_merged.csv")
hourly_steps <- read.csv("hourlySteps_merged.csv")
min_calories_narrow <- read.csv("minuteCaloriesNarrow_merged.csv")
min_calories_wide <- read.csv("minuteCaloriesWide_merged.csv")
min_intensities_narrow <- read.csv("minuteIntensitiesNarrow_merged.csv")
min_intensities_wide <- read.csv("minuteIntensitiesWide_merged.csv")
min_mets_narrow <- read.csv("minuteMETsNarrow_merged.csv")
min_sleep <- read.csv("minuteSleep_merged.csv")
min_steps_narrow <- read.csv("minuteStepsNarrow_merged.csv")
min_steps_wide <- read.csv("minuteStepsWide_merged.csv")
sleep_day <- read.csv("sleepDay_merged.csv")
weight_loginfo <- read.csv("weightLoginfo_merged.csv")
```

#### 3.3 Read all files
  
After reading all the files, as we are aiming to find the trends, we will focus on the following data:
    * Daily Activity
    * Daily Calories
    * Hourly Steps
    * Sleep Day

#### 3.4 Understanding the structure of the data
```{r }
str(daily_activity)
str(daily_calories)
str(hourly_steps)
str(sleep_day)
```

## 4. Process
#### 4.1 Number of users
```{r }
n_unique(daily_activity$Id)
n_unique(daily_calories$Id)
n_unique(hourly_steps$Id)
n_unique(sleep_day$Id)
```
<br>
Data Limitations: Based on the result, we can know that this is a very small data sample, we only have 33 samples. 

#### 4.2 Duplicates and N/As
```{r }
sum(duplicated(daily_activity))
sum(duplicated(daily_calories))
sum(duplicated(hourly_steps))
sum(duplicated(sleep_day))

which(is.na(daily_activity))
which(is.na(daily_calories))
which(is.na(hourly_steps))
which(is.na(sleep_day))
```
<br>
The results shows that sleep_day has duplicated data. No tables have N/A values.

#### 4.3 Remove duplicated data and check
```{r }
sleep_day <- sleep_day %>%
  distinct()
sum(duplicated(sleep_day))
```

#### 4.4 Rename columns
We want to ensure that all column names are using the same format in all datasets since we will merge them together. We are changing the format of all columns to lower case.
```{r Rename columns}
daily_activity <- clean_names(daily_activity)
daily_calories <- clean_names(daily_calories)
hourly_steps <- clean_names(hourly_steps)
sleep_day <- clean_names(sleep_day)
```

#### 4.5 Changing format
We will be merging the daily_activity and sleep_day tables. In this way, we can check if there is any correlation between sleep time and the total calories and total steps.

Before merging the tables, the date format is still character format, we need to change it to date format.
```{r chr to date formate}
daily_activity <- daily_activity %>%
  rename(date = activity_date) %>%
  mutate(date = as_date(date, format = "%m/%d/%Y"))

sleep_day <- sleep_day %>%
  rename(date = sleep_day) %>%
  mutate(date = as_date(date,format ="%m/%d/%Y %I:%M:%S %p" , tz=Sys.timezone()))
  ## as_date: This line converts the date column to the Date class using the as_date() function.

hourly_steps<- hourly_steps %>% 
  rename(date_time = activity_hour) %>% 
  mutate(date_time = as.POSIXct(date_time,format ="%m/%d/%Y %I:%M:%S %p" , tz=Sys.timezone()))
  ## This line converts the date_time column to the POSIXct class, which represents date and time information, using the as.POSIXct() function.
```

#### 4.6 merge tables
```{r}
daily_activity_sleep <- merge(daily_activity, sleep_day, by=c ("id", "date"))
glimpse(daily_activity_sleep)
```

## 5. Analyze

+ We need to find: 
  + Correlation between Total Steps and Sleep Time
  + Correlation between Total Steps and Calories Burned
  + Most active days in a week (-> Act)
  + Daily steps per week (find the suggested steps per day for women -> Act)
  + Daily sleep minutes per week (find the suggested sleep hours for women -> Act)
  + Usage of smart devices

#### 5.1 Correlation between Total Steps and Sleep Time
```{r}
corr_step_sleep <- cor(daily_activity_sleep$total_steps, daily_activity_sleep$total_minutes_asleep)

ggplot(daily_activity_sleep, aes(x=total_steps, y=total_minutes_asleep))+
  geom_jitter() +
  geom_smooth(color = "red") + 
  ggtitle(paste("Daily Steps vs Minutes Asleep\nCorrelation:", round(corr_step_sleep, 2))) +
  ## ggtitle() set the plot title, and the paste() function is used to create a formatted title string that includes the correlation coefficient. 
  labs(x = "Daily steps", y = "Minutes asleep") +
  theme(panel.background = element_blank(), 
        plot.title = element_text(size=14, hjust = 0.5))
        ## set the hjust (horizontal justification) argument to 0.5 to center the title horizontally within the plot area.
```
<br>
The correlation between Daily Steps and Minutes asleep is -0.19. It indicating little to no correlation between steps and amount of minutes users sleep per day.

#### 5.2 Correlation between Total Steps and Calories Burned
```{r }
corr_step_calories <- cor(daily_activity_sleep$total_steps, daily_activity_sleep$calories)

ggplot(daily_activity_sleep, aes(x=total_steps, y=calories))+
  geom_jitter() +
  geom_smooth(color="red") +
  ggtitle(paste("Daily Steps vs Calories Burned\nCorrelation:", round(corr_step_calories, 2))) +
  labs(x="Dialy steps", y="Calories Burned") +
  theme(panel.background = element_blank(), 
        plot.title = element_text(size=14, hjust = 0.5))
```
<br>
The correlation between Daily Steps and Calories Burned is 0.41. It indicating there is a strong positive relationship between them. The more steps taken, the more calories one may burn. 


#### 5.3 Check the daily average by user
```{r }
daily_average <- daily_activity_sleep %>%
  group_by(id) %>%
  summarise (mean_daily_steps = mean(total_steps), mean_daily_calories = mean(calories), mean_daily_sleep = mean(total_minutes_asleep))

head(daily_average)
```

#### 5.4 Classify active type
```{r }
active_type <- daily_average %>%
  mutate(active_type = case_when(
    mean_daily_steps < 5000 ~ "sedentary",
    mean_daily_steps >= 5000 & mean_daily_steps < 7499 ~ "lightly active", 
    mean_daily_steps >= 7500 & mean_daily_steps < 9999 ~ "fairly active", 
    mean_daily_steps >= 10000 ~ "very active"
  ))

head(active_type)
```

#### 5.5 The proportion of each type
```{r }
user_type_sum <- active_type %>%
  group_by(active_type) %>%
  summarise(total= n()) %>%
  mutate(total_percent= scales::percent (total/sum(total))) %>%
  arrange(match(active_type, c("very active", "fairly active", "lightly active", "sedentary")))
  ## use arrange(match()) to order the way you want

head(user_type_sum)
```

#### 5.6 Usage of smart device
```{r }

custom_colors <- c("#006633","#2ECC71", "#00e673", "#80ffbf")

ggplot(user_type_sum, aes(x=" ", y=total_percent, fill=active_type)) + 
  geom_bar(stat="identity", width=1) +
  coord_polar(theta="y") + 
  theme_minimal() +
  theme(axis.title.x= element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank(), 
        panel.grid = element_blank(), 
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0.5, size=14, face = "bold")) + 
  geom_text(aes(label = total_percent),
            position = position_stack(vjust = 0.5)) +
  ggtitle("Usage of Smart Device") + 
  scale_fill_manual(values = custom_colors)

```

#### 5.7 Hourly steps throughout the day
```{r}
hourly_steps <- hourly_steps %>%
  separate(date_time, into = c("date", "time"), sep= " ") %>%
  mutate(date = ymd(date)) %>%
  group_by(time) %>%
  summarize(average_steps = mean(step_total))
  
head(hourly_steps)
```

```{r }
ggplot(hourly_steps, aes(x=time, y = average_steps, fill = average_steps)) +
  geom_col()  +
  labs(title = "Hourly steps throughout the day", x="Time", y="Average Steps") + 
  scale_fill_gradient(low = "green", high = "red")+
  theme(axis.text.x = element_text(angle = 90), plot.title = element_text(size=14, hjust=0.5))
```
<br>
As shown in the plot, we can see that users are more active during lunch time from 12pm to 2pm and evenings from 5pm and 7pm.



## 6. Conclusion

1. There is no correlation between Daily Steps and Minutes Asleep.
2. There is a positive correlation between Daily Steps and Calories Burned, meaning that the more steps taken, the more calories one may burn. 
3. 42% of the users (10 users) didn't reach at least daily recommended steps by CDC, wich is 7,500 steps. 
4. Users are more active during lunch time from 12pm to 2pm and evenings from 5pm and 7pm.

Recommendations:
<br>

1. The application could provide valuable insights into the health benefits of walk exceeding 7,500 steps each day, enhancing users' understanding of the positive impact on their well-being. Aiding users in setting personalized goals according to their profiles, lifestyles, and health conditions. The intensity can gradually increase on a monthly basis. Additionally, the app sends reminders to users when they fall short of their goals.

2. The app introduces a reward system to motivate users to achieve their goals.

3. While there may be no direct correlation between daily step counts and minutes of sleep, inadequate sleep can impact users' health and indirectly influence their daily step count. The app can also send sleep reminders to users.

4. Given that the majority of our users are white-collar workers, a reasonable hourly step goal during midday from 12 pm to 2 pm and in the evenings from 5 pm to 7 pm has been set.
5. Since we encourage users to wear our product throughout the day for comprehensive data collection, long-lasting battery life is crucial. Additionally, water resistance and stylish, elegant designs are important features for our product.